\documentclass[a4paper,10pt]{report}

% Packages
\usepackage{geometry}
\usepackage{makeidx}
\usepackage{srcltx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{array}
\usepackage{wrapfig}
\usepackage{textcomp}
\usepackage{hyperref}

% Setup margins
\geometry{a4paper,left=2.5cm,right=2.5cm,top=2cm,bottom=2cm,includeheadfoot}

% Create index
\makeindex

\begin{document}

% New List Environment
\newenvironment{optlist}[1]
  {
    \begin{flushleft}\textbf{#1}\end{flushleft}
    \begin{list}{}{}
  }
  { \end{list} }

% User-Defined Commands
\newcommand{\progname}{{\sffamily Aten}}
\newcommand{\qte}[1]{\lq{}#1\rq{}}
\newcommand{\ttqte}[1]{\lq{}{\ttfamily#1}\rq{}}
\newcommand{\dqte}[1]{``#1''}
\newcommand{\its}{\hspace{20cm}}

\title{\progname{} version 0.94a \\ User's Guide}
\author{T. Youngs}
\maketitle

\tableofcontents

\clearpage

\section{Introduction}

\progname{} is first and foremost an application to ease the creation of coordinates suitable for use as the starting configurations for molecular simulation codes, ranging from single molecules \textit{in vacuo} to condensed-phase systems such as gases or liquids. Forcefield methods can be applied, for example, to minimise the energy of systems with respect to their coordinates and create \qte{good guesses} for initial configurations as input to \textit{ab initio} codes. More so, multi-molecule models can be created with Monte Carlo techniques to provide periodic systems for use in molecular dynamics.\\

For the computional chemist, it is often the case that coordinates (new or existing) need to be edited in a specific way to create the configuration necessary to perform the job at hand, and in the opinion of the author, this is a job not handled well by current the current stock of free codes available. \progname{} was initially created as a method of learning various programming, visual, and algorithmic techniques, but has grown into an application suitable for use in many situations by the simulation community.\\

Comments and suggestions on the facilities available (or not) in the code are welcome and will be warmly accepted. If a feature is missing or inadequate and of general applicability, I will most likely include it in the program. Please don't hesitate to write to me at the address provided below. In particular, if you take the time to write a file filter for use with your favourite code, or a definitions file for a forcefield, I would be happy to include as part of the distribution.\\

\progname{} is released under the GPL, meaning you may copy, change, and distribute it as you see fit, provided you acknowledge the original copyright. If, however, you have a burning desire to improve or extend the code in some way, and wish to do so under your own steam, I would be happy to consider including your modifications into the official stream. Contact me to discuss your brilliance at interpreting my code!\\

Tristan Youngs\\
9$^\textrm{th}$ November 2007\\
t.youngs@qub.ac.uk\\

% Include sections
\include{install}
\include{quickstart}
\include{features}
\include{cli}
\include{gui}
\include{expression}
\include{typing}
\include{ff}
\include{filtershandbook}
\include{scriptorshandbook}

% Create index
\printindex

\end{document}
